---
title: "Pseudotime_Generation_vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pseudotime_Generation_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Example Generation of Pseudotime Results for LineageDE:

```{r setup}
library(LineageDE)
```

Here we utilize a modified version of samplePseudotime (samplePseudotime_HSPC) that generates results like those in ExampleInputData provided with LineageDE

The function samplePseudotime can be similarly utilized to generate your own pseudotime data (and samples) utilizing Slingshot (Street et al. BMC Genomics 2018)

samplePseudotime can be adapted for other pseudotime inference algorithms

## Load in Data Inputs
```{r}
data("exPseudotimeInput")
```
This input Data contains HSPCs (hematopoietic stem and progenitor cells) from two conditions (WT & IkBKO) and two replicates (Rep1 & Rep2).  (Data subsampled from publication Singh et al.)

Already provided are basic inputs for Slingshot:
cellData = PCA (or other dim reduction) coordinates
clusters (utilized by slingshot) that specifies to which cluster cells belong to
(starting and ending clusters will be provided to slingshot)

Note: to take advantage of parallelization along samples and if RhpcBLASctl installed,
run the following command to avoid multithreaded operations:
```{r}
#RhpcBLASctl::blas_set_num_threads(1)
```

## Generate Pseudotime Results
```{r}
pseudoResults = samplePseudotime_HSPC(exPseudotimeInput$cellData, exPseudotimeInput$conditions, exPseudotimeInput$clusters, start.cluster="HSC", end.clusters=c("ErP", "MkP","MyP", "LyP"), experiments=exPseudotimeInput$experiments, nSamples=10, pseudo.cond = "WT")
```

Here we specified conditions (genotypes) and experiments (replicates).
The starting cluster for slingshot is "HSC".
The ending clusters are "ErP", "MkP","MyP", and "LyP".
We specify pseudo.cond = "WT" so that only wildtype cells are utilized to generate the pseudotime trajectories (all cells in cellData will be projected onto these trajectories and used as input for LineageDE).
nSamples should be increased (default 100) and calculations can be accelerated by setting parallel = TRUE, n_cores = # (utilizes BiocParallel)

Returned are original and sampled pseudotime results than can be utilized by LineageDE.


